\RequirePackage{hyperref}%%%
%%% The acronyms.
%%%
%
\newacronym[description={%
An \emph{Application Programming Interface} is a set of rules or protocols that enables one software application or component to use or communicate with another~\cite{G2024WIAA}.%
}]{API}{API}{Application Programming Interface}%
\newacronym[description={Artificial Intelligence, see, e.g.,~\cite{RN2022AIAMA}}]{AI}{AI}{Artificial Intelligence}%
%
\newacronym[description={%
\emph{Comma-Separated Values} is a very common and simple text format for exchanging tabular or matrix data~\cite{RFC4180}. %
Each row in the text file represents one row in the table or matrix. %
The elements in the row are separated by a fixed delimiter, usually a comma~(\inQuotes{,}), sometimes a semicolon~(\inQuotes{;}). %
\python\ offers some out-of-the-box CSV support in the \pythonilIdx{csv}~module~\cite{PSF2024CCFRAW}.%
},sort={CSV}]{csv}{CSV}{Comma-Separated Values}%
%
\newacronym[description={%
A \emph{database} is an organized collection of structured information or data, typically stored electronically in a computer system. %
Databases are discussed in our book \citetitle{databases}~\cite{databases}.%
}]{db}{DB}{database}%
\protected\gdef\db{\pgls{db}}%%
%
\newacronym[description={%
A \emph{database administrator} is the person or group responsible for the effective use of database technology in an organization or enterprise.%
}]{dba}{DBA}{database administrator}%
%
\newacronym[plural={DBMSes},description={%
A \emph{database management system} is the software layer located between the user or application and the~\pgls{db}. %
The DBMS allows the user/application to create, read, write, update, delete, and otherwise manipulate the data in the~\pgls{db}~\cite{YM2024DDMSD}.%%
}]{dbms}{DBMS}{database management system}%
\protected\gdef\db{\glslink{dbms}}%%
%
\newacronym[plural={DBSes},description={%
A \emph{database system} is the combination of a \pgls{db} and a the corresponding \glslink{dbms}, i.e., basically, an installation of a \glslink{dbms} on a computer together with one (or multiple) \pglspl{db}. %
DBS = \acrshort{db}~+~\acrshort{dbms}.%
}]{dbs}{DBS}{database system}%
%
\newacronym[plural={DSes},description={Data Science, see, e.g.,~\cite{G2019DSFSFPWP}.}]{DS}{DS}{Data Science}%
\newacronym{ERD}{ERD}{Entity Relationship Diagram.}%
\newacronym[plural={GIGO},description={Garbage In--Garbage Out, see, e.g.,~\cite{PKBCBR2005GIGOAELAOMBEUP}}]{GIGO}{GIGO}{Garbage In--Garbage Out.}%
\newacronym{GUI}{GUI}{Graphical User Interface.}%
\newacronym{HR}{HR}{Human Resources Department.}%
\newacronym[description={Hyper Text Transfer Protocol, the protocol linking web browsers to web servers in the \pgls{WWW}~\cite{BL1991RQOHL,RFC1945,RFC2068,RFC9110,G2017HP}.}]{HTTP}{HTTP}{Hyper Text Transfer Protocol}%
\newacronym[description={Hypertext Transfer Protocol Secure, i.e., an encrypted variant of~\pgls{HTTP} where data is sent over~\pgls{TLS}~\cite{RFC9110,S2021SCG2HTH}.}]{HTTPS}{HTTPS}{Hypertext Transfer Protocol Secure}%
%
\newacronym[description={%
An \emph{Integrated Developer Environment} is a program that allows the user do multiple different activities required for software development in one single system. %
It often offers functionality such as editing source code, debugging, testing, or interaction with a distributed version control system. %
For \python, we recommend using \pycharm.%
}]{ide}{IDE}{Integrated Development Environment.}%
%
\newacronym[plural={JSON},description={%
\emph{JavaScript Object Notation} is a data interchange format~\cite{E2017SE4TJDIS,RFC8259} based on \pgls{javascript}~\cite{E1999ELS} syntax.%
},sort={JSON}]{json}{JSON}{JavaScript Object Notation}%
%
\newacronym[plural={ML},description={Machine Learning, see, e.g.,~\cite{SSBD2014UMLFTTA}}]{ML}{ML}{Machine Learning.}%
\newacronym[plural={OOP}]{OOP}{OOP}{Object-Oriented Programming.}%
\newacronym[plural={OSes},description={Operating System, the system that runs your computer, see, e.g., \linux\ and \windows.}]{OS}{OS}{Operating System}%
%
\newacronym[plural={regexes},description={%
A \emph{Regular Expression}, often called \inQuotes{regex} for short, is a sequence of characters that define a search pattern for text strings~\cite{K2024REH,PSF2024RREO,N2018RQSRUAURE,N2019AITRE}. %
In \python, the \pythonilIdx{re} module offers functionality work with regular expressions~\cite{K2024REH,PSF2024RREO}.%
}]{regex}{regex}{regular expression}%
%
\newacronym[plural={SQL},description={Structured Query Language, see \cite{C20245YOQ}.}]{SQL}{SQL}{Structured Query Language}%
\protected\gdef\sql{\pgls{SQL}}%
%
\newacronym[plural={SSH},description={Secure Shell Transport Layer Protocol, a protocol that provides users a secure way to access computers over an unsecure network~\cite{RFC4253,AP2024WWS,BSB2005STSSTDG}.}]{SSH}{SSH}{Secure Shell Transport Layer Protocol}%
%
\newacronym[description={%
The \emph{standard error stream} is one of the three pre-defined streams of a console process (together with the \gls{stdin} and the \gls{stdout})~\cite{J2024PTOGBSI8IS12ESSSSIS}. %
It is the text stream to which the process writes information about errors and exceptions. %
If an uncaught \pythonilIdx{Exception} is raised in \python\ and the program terminates, then this information is written to \pgls{stderr}. %
If you run a program in a \pgls{terminal}, then the text that a process writes to its \pgls{stderr} appears in the console.%
},sort={stderr}]{stderr}{\texttt{stderr}}{standard error stream}%
%
\newacronym[description={%
The \emph{standard input stream} is one of the three pre-defined streams of a console process (together with the \pgls{stdout} and the \pgls{stderr})~\cite{J2024PTOGBSI8IS12ESSSSIS}. %
It is the text stream from which the process reads its input text, if any. %
The \python\ instruction \pythonilIdx{input} reads from this stream. %
If you run a program in a \pgls{terminal}, then the text that you type into the terminal while the process is running appears in this stream.%
},sort={stdin}]{stdin}{\texttt{stdin}}{standard input stream}%
%
\newacronym[description={%
The \emph{standard output stream} is one of the three pre-defined streams of a console process (together with the \pgls{stdin} and the \pgls{stderr})~\cite{J2024PTOGBSI8IS12ESSSSIS}. %
It is the text stream to which the process writes its normal output. %
The \pythonilIdx{print} instruction of \python\ writes text to this stream. %
If you run a program in a \pgls{terminal}, then the text that a process writes to its \pgls{stdout} appears in the console.%
},sort={stdout}]{stdout}{\texttt{stdout}}{standard output stream}%
%
\newacronym[description={Transport Layer Security, a protol for encrypted communication over the internet~\cite{RFC8446,DS2024TCID,S2021SCG2HTH}, used by, e.g., \pgls{HTTPS}.}]{TLS}{TLS}{Transport Layer Security}%
%
\newacronym[plural={UCS},description={Universal Coded Character Set, see \gls{unicode}}]{UCS}{UCS}{Universal Coded Character Set.}%
%
\newacronym[description={Uniform Resource Locator, identifiers for resources in the \pgls{WWW}, most notable the text you write into web browsers to visit websites~\cite{RFC1738}.}]{URL}{URL}{Uniform Resource Locator}%
%
\newacronym[plural={UTF8},description={%
The \emph{\acrshort{UCS}~Transformation Format~8} is one standard for encoding \pgls{unicode} characters into a binary format that can be stored in files~\cite{RFC3629,ISOIEC106462020ITUCCSU}. %
It is the world wide web's most commonly used character encoding, where each character is represented by one to four bytes. %
It is backwards compatible with ASCII\optionalRef{sec:unicodeChars}{~(see~\cref{sec:unicodeChars})}.%
},sort={UTF-8}]{utf8}{UTF\nobreakdashes-8}{\acrshort{UCS}~Transformation Format~8}%
%
\newacronym[plural={VCSes},description={%
A \emph{Version Control System} is a software which allows you to manage and preserve the historical development of your program code~\cite{T2024BGAGVCPMATFTND}. %
A distributed VCS allows multiple users to work on the same code and upload their changes to the server, which then preserves the change history.
The most popular distributed VCS is \pgls{git}. %
}]{vcs}{VCS}{Version Control Systems}%
%
\newacronym[description={World Wide Web~\cite{BL1991RQOHL,DDD2011IWWWHTP}}]{WWW}{WWW}{World Wide Web.}%
%
\newacronym[plural={XML},description={%
\sloppy%
The \emph{Extensible Markup Language} is a text-based language for storing and transporting of data~\cite{BPSMM2008EMLX1FE,K2019ITXJY,CH2013XFCAMLTMC}. %
It allows you to define elements in the form \xmlil{<myElement myAttr="x">...text..</myElement>}. %
Different from \pgls{csv}, elements in XML can be hierarchically nested, like~\textil{<a><b><c>test</c></b><b>bla</b></a>}, and thus easily represent tree structures. %
XML is one of most-used data interchange formats. %
To process XML in \python, use the \pythonilIdx{defusedxml} library~\cite{H2021D0XBPFPSM}, as it protects against several security issues.%
},sort={XML}]{xml}{XML}{Extensible Markup Language}%
%
\newacronym[plural={YAML},description={%
\emph{{YAML}~Ain't Markup Language\texttrademark} is a human-friendly data serialization language for all programming languages~\cite{DNMAASBE2021YAMLYV1,K2019ITXJY,CGTYB2022YFFDCAIE}. %
It is widely used for configuration files in the DevOps environment. %
See \url{https://yaml.org} for more information.%
},sort={YAML}]{yaml}{YAML}{{YAML}~Ain't Markup Language\texttrademark}%
%
